---
title: "AVC_BCO_bone_nondup_genomic_analysis"
author: "Max Cummins"
date: "`r Sys.Date()`"
output:
    html_document:
        theme: spacelab
        toc: true
        toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
#Read in our packages
library(tidyverse)
library(abricateR)
library(dplyr)
library(ggplot2)
library(ggtree)
library(pheatmap)
library(phytools)
library(plotly)
library(readr)
library(readxl)
library(paletteer)

if("MASS" %in% (.packages())){
  detach("package:MASS", unload=TRUE) 
}

#Define our not in function
`%nin%` <- Negate(`%in%`)

```

## Outline

## File summary

A GitHub repository available at
<https://github.com/maxlcummins/bonechook> contains all scripts and
files required to rerun analysis detailed in this document.

Below are the files associated with this research project and report.

**Insert Filetree**

### Genomic data

Genomic data was generated using snakemake using the following [snakefile](https://github.com/maxlcummins/pipelord/tree/main/workflow/Snakefile).

```{bash run_genomics_workflow, eval=FALSE}
#Run our genomics workflow
nohup snakemake --use-conda --conda-frontend mamba -T 10 --rerun-incomplete -s workflow/Snakefile --configfile config/AVC_FH.yaml -j 36 -k -p > /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/AVC_FH_Genotype_all.out.log 2> /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/AVC_FH_Genotype_all.err.log &

# Jupiter6
# JobID = 41101
# Fri Jun  3 22:38:03 AEST 2022
```

### Mashtree

We also generated a mashtree in our QC step, but we will now generate another with only the ones which passed QC.

```{bash run_mashtree, eval=FALSE}
#Deactivate our old environment
source deactivate

#Activate our new environment
source activate mashtree

#Run mashtree
nohup mashtree --mindepth 0 --tempdir /tmp/mashtree_QC_pass --numcpus 20 --outmatrix /projects/AusGEM/Users/Max/AVC_FH/analysis/mashtree/mashtree_QC_pass.matrix --outtree /projects/AusGEM/Users/Max/AVC_FH/analysis/mashtree/mashtree_QC_pass.tree /projects/AusGEM/Users/Max/AVC_FH/data/symlink_QC_pass/*.R1.fastq.gz > /projects/AusGEM/Users/Max/AVC_FH/logs/mashtree/mashtree_QC_pass.out.log 2> /projects/AusGEM/Users/Max/AVC_FH/logs/mashtree/mashtree_QC_pass.err.log &

# Jupiter8
# JobID = 91625
# Mon Jun  6 11:52:44 AEST 2022
```

Now we generate the tree consisting of the BCO samples we filtered for in our QC
script. These samples are different from the full BCO cohort as they:

* Are limited to one sample per bird
* Are only sourced from bone sites

```{bash run_mashtree2, eval=FALSE}
#Deactivate our old environment
source deactivate

#Activate our new environment
source activate mashtree

#Run mashtree
nohup mashtree --mindepth 0 --tempdir /tmp/mashtree_QC_pass_AVC_BCO_bone_nondup --numcpus 6 --outmatrix /projects/AusGEM/Users/Max/AVC_FH/analysis/mashtree/mashtree_QC_pass_AVC_BCO_bone_nondup.matrix --outtree /projects/AusGEM/Users/Max/AVC_FH/analysis/mashtree/mashtree_QC_pass_AVC_BCO_bone_nondup.tree /projects/AusGEM/Users/Max/AVC_FH/data/sampled_bonesites_and_AVC/*.R1.fastq.gz > /projects/AusGEM/Users/Max/AVC_FH/logs/mashtree/mashtree_QC_pass_AVC_BCO_bone_nondup.out.log 2> /projects/AusGEM/Users/Max/AVC_FH/logs/mashtree/mashtree_QC_pass_AVC_BCO_bone_nondup.err.log &

# Jupiter5
# JobID = 245746
# Tue Jun  7 13:58:55 AEST 2022
```

Now we run our pangenomics workflow to generate our phylogeny and pangenome for our bone nondupe cohort (with AVC).

```{bash run_pangenomics_workflow, eval=FALSE}
#Run our pangenomics/phylogenetic workflow
nohup snakemake --use-conda --conda-frontend mamba -T 10 --rerun-incomplete -s workflow/Treebuild_panaroo.smk --configfile config/AVC_FH.yaml -j 30 -k -p > /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/AVC_FH_pangenome_bone_nondup.out.log 2> /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/AVC_FH_pangenome_bone_nondup.err.log &

# Jupiter12
# JobID = 35408
# Wed Jun 15 09:41:44 AEST 2022
```

```{bash run_pangenomics_workflow2, eval=FALSE}
#Run our pangenomics/phylogenetic workflow
nohup snakemake --use-conda --conda-frontend mamba --rerun-incomplete -s workflow/Treebuild_panaroo.smk --configfile config/AVC_FH.yaml -j 30 -k -p > /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/AVC_FH_pangenome_bone_nondup.out2.log 2> /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/AVC_FH_pangenome_bone_nondup.err2.log &

# Jupiter15
# JobID = 43678
# Thu Jun 16 09:56:35 AEST 2022
```

Now we run it again for our FH isolates, and FH ST117, ST95, ST69 and ST57
```{bash run_pangenomics_workflow3, eval=FALSE}
#Run our pangenomics/phylogenetic workflow
nohup snakemake --use-conda --conda-frontend mamba --rerun-incomplete -s workflow/Treebuild_panaroo.smk --configfile config/AVC_FH.yaml -j 30 -k -p > /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/FH_only.out.log 2> /projects/AusGEM/Users/Max/AVC_FH/logs/snakemake/FH_only.err.log &

# Jupiter10
# JobID = 19994
# Mon Jul  4 11:07:48 AEST 2022
```


## Genomic analysis

After downloading all of the resulting files from the remote server we
are ready to begin our QC.

```{r read_in_file_paths, echo=FALSE, tidy = TRUE}
#Load paths to files needed for the script
#Assembly stats file
assembly_stats_path <-
        file.path("analysis/assembly_stats.txt")
#A list of taxonomic IDs and their respective scientific names
tax_id_converter_path  <-
        file.path("delims/taxid_to_scientific_name.txt")
#A list of expected genome sizes.
genome_sizes_path  <-
        file.path("delims/species_genome_size.txt")

abricate_path <- file.path("analysis/genotype.txt")
pointfinder_path <- file.path("analysis/pointfinder.txt")
pMLST_path <- file.path("analysis/pMLST.txt")
serotype_path <- file.path("analysis/ectyper.txt")
fimtype_path <- file.path("analysis/fimtyper.txt")
phylogroup1_path <- file.path("analysis/ezclermont.txt")
phylogroup2_path <- file.path("analysis/clermontyping.txt")
mlst_path <- file.path("analysis/mlst.txt")
bracken_path <- file.path("analysis/bracken_report.txt")
checkm_path <- file.path("analysis/trees/checkm_qa.txt")
metadata_path <- file.path("delims/merged_metadata_AVC_FH.txt")

enterobase_data_path <- file.path("delims/enterobase.txt")

tree_path <- file.path("analysis/trees/old/mashtree_QC_pass.tree")

QC_pass_path <- file.path("delims/QC_data.txt")

#Provide output names
output_name <- "BCO_APEC"
```

```{r quiet_abricate, echo=FALSE,include=FALSE}
#Read in our tree
tree <- read.tree(tree_path)

#Midpoint root our tree
tree <- midpoint.root(tree)

#Trim off excess characters from the tree tip labels
tree$tip.label <- gsub(".R1","",tree$tip.label)

#Run AbricateR to summarise genomic data
abricateR::abricateR(
        abricate_in = abricate_path,
        output = output_name,
        output_directory = output_dir,
        identity = 90,
        length = 90,
        writecsv = FALSE,
        pointfinder_data = pointfinder_path#,
        #pMLST_data = pMLST_path
)

#Filter our collection
BCO_APEC_simple_summary_N90L90 <- BCO_APEC_simple_summary_N90L90 %>% filter(name %in% tree$tip.label)
BCO_APEC.N90.L90.PASS <- BCO_APEC.N90.L90.PASS %>% filter(name %in% tree$tip.label)
BCO_APEC.N90.L90.FAIL <- BCO_APEC.N90.L90.FAIL %>% filter(name %in% tree$tip.label)
BCO_APEC_co_occurence_N90L90 <- BCO_APEC_co_occurence_N90L90 %>% filter(name %in% tree$tip.label)

#Read in our filtered metadata
metadata <- read_delim(metadata_path, delim = "\t", show_col_types = FALSE)

#Rename our name column
metadata <- metadata %>% rename(name= `Sample Name`)

#Filter our metadata to our tree
metadata <- metadata %>% filter(name %in% tree$tip.label)

#Bind our genotypic data and metadata
genometa <- left_join(metadata, BCO_APEC_simple_summary_N90L90, by = "name") %>% as.data.frame()

#Move our working name to the start
genometa <- genometa %>% dplyr::select(name, everything())

#Apply our names as rownames
rownames(genometa) <- genometa$name

#Read in our MLST data
mlst <- read_delim(mlst_path, delim = "\t", show_col_types = FALSE)

#Bind our MLST data
genometa <- left_join(genometa, mlst, by = "name")

#categorise top ten STs or other
genometa <- genometa %>% mutate(ST_other = fct_lump(ST, n = 10))

#Read in our fim data
fimtype <- read_delim(fimtype_path, delim = "\t", show_col_types = FALSE)

#Remove potential duplicate hits, only take the first.
#AVC160 appears to have a duplicated fimH gene (same type, same scaffold)
fimtype <- fimtype %>% dplyr::select(name, Fimtype) %>% unique()

#Bind our fimtype data
genometa <- left_join(genometa, fimtype, by = "name")

#Read in our serotype data
serotype <- read_delim(serotype_path, delim = "\t", show_col_types = FALSE)

#Rename name column
serotype <- serotype %>% rename('name' = 'Name')

#Bind our serotype data
genometa <- left_join(genometa, serotype, by = "name")

#Read in our ezclermont data
ezclermont <- read_delim(phylogroup1_path, delim = "\t", show_col_types = FALSE)

#Remove spaces at start of names for some reason
ezclermont <- ezclermont %>% mutate(name = str_replace(name, '^ ', ''))

#Bind our ezclermont data
genometa <- left_join(genometa, ezclermont, by = "name")

#Read in our clermontyper data
clermontyper <- read_delim(phylogroup2_path, delim = "\t", show_col_types = FALSE)

#Bind our ezclermont data
genometa <- left_join(genometa, clermontyper, by = "name")


```

## Collection breakdown

We will start by producing some figures relating to the collection composition:

### Breakdown by year
```{r year_breakdown, echo=FALSE}
library(ggplot2)
library(magrittr)

df2 <- metadata %>% group_by(collection_year) %>% summarise(counts = n(), .groups = "keep")

#ggplot(df2, aes(x=collection_year, y=counts, fill=Collection)) +
#  geom_bar() + theme(axis.text.x = element_text(angle=90)) + xlab("Collection Year") + ylab("Count")

metadata$collection_year[metadata$collection_year == "NA"] <- NA

years <- min(unique(sort(metadata$collection_year))):max(unique(sort(metadata$collection_year)))

cols_fun <- colorRampPalette(c("blue", "red"))
year_cols <- cols_fun(length(years))

names(year_cols) <- years

ylimit <- max(df2$counts) + (max(df2$counts) * 0.1)

#Treemap
year_dist_plot <- ggplot(metadata, aes(x = collection_year, fill = as.character(collection_year))) +
        geom_bar(stat = "count") +
        scale_fill_manual(
                name = "Collection Year",
                aesthetics = c("colour", "fill"),
                values = year_cols,
                na.value = 'grey'
        ) +
        theme(axis.text.x = element_text(angle = 90)) +
        xlab("Collection Year") +
        ylab("Count") +
        scale_y_continuous(expand = c(0, 0),
                           limits = c(0, ylimit)
                           ) +
        scale_x_continuous(expand = c(0, 0),
                           limits = c(min(years)-1, max(years)+1),
                           labels = (min(years)-1):(max(years)+1),
                           breaks = (min(years)-1):(max(years)+1)
                           ) 

year_dist_plot
```

### Generate our first tree

```{r base_tree, echo=FALSE}


#Create a list of ST other names
ST_other_names <- genometa %>% group_by(ST_other) %>% tally(sort = TRUE) %>% pull(ST_other)

#Generate a list of colours for STs
ST_other_cols <- c(#Red
"#e6194B",
#Green
"#3cb44b",
##Yellow
#"#ffe119",
#Blue
"#4363d8",
#Orange
"#f58231",
#Purple
"#911eb4",
#Cyan
"#42d4f4",
#Magenta
"#f032e6",
#Lime
"#bfef45",
#Pink
"#fabed4",
#Teal
"#469990",
##Lavender
#"#dcbeff",
#Brown
"#9A6324",
##Beige
#"#fffac8",
#Maroon
"#800000",
#Mint
"#aaffc3",
#Olive
"#808000",
#Apricot
"#ffd8b1",
#Navy
"#000075",
#Grey
"#a9a9a9",
##White
#"#ffffff",
#Black
"#000000")

#Trim our list of colours for STs
ST_other_cols <- ST_other_cols [1:length(ST_other_names)]

#Name our ST colours
names(ST_other_cols) <- ST_other_names

#Define our collection names
collection_names <- c("AVC", "BCO")

#Define our collection colours
collection_colours <- c("black", "red")

#Assign names to our collection colours
names(collection_colours) <- collection_names

#Bind our list of colours 
mycolslist <- c(collection_colours, ST_other_cols)

#Get rid of ugly complex O types
genometa$`O-type` <- gsub('/O', '/', genometa$`O-type`)

#Build tree
p1 <- ggtree(tree, layout = 'circular') %<+%
        genometa +
        geom_tiplab(size = 1.5,
                    align = TRUE,
                    offset = 0.002,
                    linesize = 0.15,
                    aes(color = as.factor(ST))
                ) +
        geom_tiplab(size = 1.5,
                    align = TRUE,
                    linetype = NULL,
                    offset = 0.007,
                    aes(label = paste0("ST",as.character(ST),":", `O-type`,":", `H-type`))
                ) +
        geom_tiplab(size = 1.5,
                    align = TRUE,
                    linetype = NULL,
                    offset = 0.018,
                    aes(label = Clermont_type)
                ) +
        geom_tippoint(size = 0.5,
                      aes(colour = Collection)
                ) +
        scale_fill_manual(
                        aesthetics = c("colour", "fill"),
                        values = mycolslist,
                        na.value = 'grey'
                        ) 

ST_coord <- fortify(p1$data) %>%
        filter(isTip == TRUE) %>%
        dplyr::select(label, ST_other, y) %>%
        arrange(y) %>%
        group_by(ST_other) %>%
        summarise(
                firstocc = first(label),
                lastocc = last(label),
                count = n()
        )


#Bind our colours and names
STcols_df <- suppressWarnings(as.data.frame(cbind(ST_other_cols, as.character(ST_other_names))))

#Rename our coloumns
STcols_df <- STcols_df %>% rename(color = ST_other_cols, ST_other = V2)

#Join our df of coordinates for our annotations with our colours for our annotations
ST_coord <- left_join(ST_coord, STcols_df, by = "ST_other")



find_firstlast <- function(x, y){
        ST_other_coord <- fortify(p1$data) %>%
        filter(isTip == TRUE) %>%
        dplyr::select(label, Clermont_type, y, angle) %>%
        arrange(y) %>%
        filter(Clermont_type == x) %>%
        summarise(
                firstocc = first(label),
                lastocc = last(label),
                angle = (median(angle))
        )
                
        #Dynamically set hjust based on angle of label
        ST_other_coord <- ST_other_coord %>% mutate(hjust = if_else(angle > 90 & angle < 270, 1, 0))
        
        #Modify angle for labels beyong 90 degrees and before 270 so they arent upside down
        ST_other_coord <- ST_other_coord %>% mutate(angle = if_else(angle > 90 & angle < 270, angle + 180, angle))
        #Angles cant be over 360 (i think) so lets reduce by 360 if they are
        ST_other_coord <- ST_other_coord %>% mutate(angle = if_else(angle > 360, angle - 360, angle))
        
        return(ST_other_coord[[y]])
}

#Define some parameters for our plotting
strip_fontsize <- 5
strip_offset <- 0.02
strip_text_offset <- 0.003
strip_barsize <- 3
strip_text_hjust <- 1

#Plot our core tree
p2 <- p1 + geom_strip(
        taxa1 = "FH83",
        taxa2 = "FH58",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[1],
        label = "A",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("A", 3),
        hjust = find_firstlast("A", 4)
) + geom_strip(
        taxa1 = "AVC94",
        taxa2 = "FH164",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[2],
        label = "B1",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("B1", 3),
        hjust = find_firstlast("B1", 4)
) + geom_strip(
        taxa1 = "FH101",
        taxa2 = "AVC183",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[3],
        label = "B2",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("B2", 3),
        hjust = find_firstlast("B2", 4)
) + geom_strip(
        taxa1 = find_firstlast("C", 1),
        taxa2 = find_firstlast("C", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[4],
        label = "C",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("C", 3),
        hjust = find_firstlast("C", 4)
) + geom_strip(
        taxa1 = "FH141",
        taxa2 = "FH22",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[5],
        label = "D",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("D", 3),
        hjust = find_firstlast("D", 4)
) + geom_strip(
        taxa1 = "FH45",
        taxa2 = "FH59",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[6],
        label = "E",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("E", 3),
        hjust = find_firstlast("E", 4)
) + 
#        geom_strip(
#        taxa1 = find_firstlast("F", 1),
#        taxa2 = find_firstlast("F", 2),
#        barsize = strip_barsize,
#        color = paletteer_d("ggsci::nrc_npg")[7],
#        label = "F",
#        fontsize = strip_fontsize,
#        offset = strip_offset,
#        offset.text = strip_text_offset,
#        angle = find_firstlast("F", 3),
#        hjust = find_firstlast("F", 4)
#) + 
        geom_strip(
        taxa1 = "FH154",
        #Manually change taxa label
        taxa2 = "AVC107",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[8],
        label = "G",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("G", 3),
        hjust = find_firstlast("G", 4)
) +
#        geom_strip(
#        taxa1 = find_firstlast("U", 1),
#        taxa2 = find_firstlast("U", 2),
#        barsize = strip_barsize,
#        color = paletteer_d("ggsci::nrc_npg")[9],
#        label = "U",
#        fontsize = strip_fontsize,
#        offset = strip_offset,
#        offset.text = strip_text_offset,
#        angle = find_firstlast("U", 3),
#        hjust = find_firstlast("U", 4)
#) +
        theme(legend.position = "none")
```

### Generate our second tree

```{r base_tree2, echo=FALSE}


#Create a list of ST other names
ST_other_names <- genometa %>% group_by(ST_other) %>% tally(sort = TRUE) %>% pull(ST_other)

#Generate a list of colours for STs
ST_other_cols <- c(#Red
"#e6194B",
#Green
"#3cb44b",
##Yellow
#"#ffe119",
#Blue
"#4363d8",
#Orange
"#f58231",
#Purple
"#911eb4",
#Cyan
"#42d4f4",
#Magenta
"#f032e6",
#Lime
"#bfef45",
#Pink
"#fabed4",
#Teal
"#469990",
##Lavender
#"#dcbeff",
#Brown
"#9A6324",
##Beige
#"#fffac8",
#Maroon
"#800000",
#Mint
"#aaffc3",
#Olive
"#808000",
#Apricot
"#ffd8b1",
#Navy
"#000075",
#Grey
"#a9a9a9",
##White
#"#ffffff",
#Black
"#000000")

#Trim our list of colours for STs
ST_other_cols <- ST_other_cols [1:length(ST_other_names)]

#Name our ST colours
names(ST_other_cols) <- ST_other_names

#Define our collection names
collection_names <- c("AVC", "BCO")

#Define our collection colours
collection_colours <- c("black", "red")

#Assign names to our collection colours
names(collection_colours) <- collection_names

#Bind our list of colours 
mycolslist <- c(collection_colours, ST_other_cols)

#Get rid of ugly complex O types
genometa$`O-type` <- gsub('/O', '/', genometa$`O-type`)

#Build tree
p1_linear <- ggtree(tree) %<+%
        genometa +
        geom_tiplab(size = .75,
                    align = TRUE,
                    offset = 0.002,
                    linesize = 0.15,
                    aes(color = as.factor(ST))
                ) +
        geom_tiplab(size = .75,
                    align = TRUE,
                    linetype = NULL,
                    offset = 0.01,
                    aes(label = paste0("ST",as.character(ST),":", `O-type`,":", `H-type`))
                ) +
        geom_tiplab(size = 0.75,
                    align = TRUE,
                    linetype = NULL,
                    offset = 0.025,
                    aes(label = Clermont_type)
                ) +
        geom_tippoint(size = 0.5,
                      aes(colour = Collection)
                ) +
        scale_fill_manual(
                        aesthetics = c("colour", "fill"),
                        values = mycolslist,
                        na.value = 'grey'
                        ) 

ST_coord <- fortify(p1_linear$data) %>%
        filter(isTip == TRUE) %>%
        dplyr::select(label, ST_other, y) %>%
        arrange(y) %>%
        group_by(ST_other) %>%
        summarise(
                firstocc = first(label),
                lastocc = last(label),
                count = n()
        )


#Bind our colours and names
STcols_df <- suppressWarnings(as.data.frame(cbind(ST_other_cols, as.character(ST_other_names))))

#Rename our coloumns
STcols_df <- STcols_df %>% rename(color = ST_other_cols, ST_other = V2)

#Join our df of coordinates for our annotations with our colours for our annotations
ST_coord <- left_join(ST_coord, STcols_df, by = "ST_other")



find_firstlast <- function(x, y){
        ST_other_coord <- fortify(p1_linear$data) %>%
        filter(isTip == TRUE) %>%
        dplyr::select(label, Clermont_type, y, angle) %>%
        arrange(y) %>%
        filter(Clermont_type == x) %>%
        summarise(
                firstocc = first(label),
                lastocc = last(label),
                angle = (median(angle))
        )
                
        #Dynamically set hjust based on angle of label
        ST_other_coord <- ST_other_coord %>% mutate(hjust = if_else(angle > 90 & angle < 270, 1, 0))
        
        #Modify angle for labels beyong 90 degrees and before 270 so they arent upside down
        ST_other_coord <- ST_other_coord %>% mutate(angle = if_else(angle > 90 & angle < 270, angle + 180, angle))
        #Angles cant be over 360 (i think) so lets reduce by 360 if they are
        ST_other_coord <- ST_other_coord %>% mutate(angle = if_else(angle > 360, angle - 360, angle))
        
        return(ST_other_coord[[y]])
}

#Define some parameters for our plotting
strip_fontsize <- 5
strip_offset <- 0.025
strip_text_offset <- 0.003
strip_barsize <- 3
strip_text_hjust <- 1

#Plot our core tree
p2_linear <- p1_linear + geom_strip(
        taxa1 = find_firstlast("A", 1),
        taxa2 = find_firstlast("A", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[1],
        label = "A",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("A", 3),
        hjust = find_firstlast("A", 4)
) + geom_strip(
        taxa1 = find_firstlast("B1", 1),
        taxa2 = find_firstlast("B1", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[2],
        label = "B1",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("B1", 3),
        hjust = find_firstlast("B1", 4)
) + geom_strip(
        taxa1 = find_firstlast("B2", 1),
        taxa2 = find_firstlast("B2", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[3],
        label = "B2",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("B2", 3),
        hjust = find_firstlast("B2", 4)
) + geom_strip(
        taxa1 = find_firstlast("C", 1),
        taxa2 = find_firstlast("C", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[4],
        label = "C",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("C", 3),
        hjust = find_firstlast("C", 4)
) + geom_strip(
        taxa1 = find_firstlast("D", 1),
        taxa2 = find_firstlast("D", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[5],
        label = "D",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("D", 3),
        hjust = find_firstlast("D", 4)
) + geom_strip(
        taxa1 = find_firstlast("E", 1),
        taxa2 = find_firstlast("E", 2),
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[6],
        label = "E",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("E", 3),
        hjust = find_firstlast("E", 4)
) + 
#        geom_strip(
#        taxa1 = find_firstlast("F", 1),
#        taxa2 = find_firstlast("F", 2),
#        barsize = strip_barsize,
#        color = paletteer_d("ggsci::nrc_npg")[7],
#        label = "F",
#        fontsize = strip_fontsize,
#        offset = strip_offset,
#        offset.text = strip_text_offset,
#        angle = find_firstlast("F", 3),
#        hjust = find_firstlast("F", 4)
#) + 
        geom_strip(
        taxa1 = find_firstlast("G", 1),
        #Manually change taxa label
        taxa2 = "FH171",
        barsize = strip_barsize,
        color = paletteer_d("ggsci::nrc_npg")[8],
        label = "G",
        fontsize = strip_fontsize,
        offset = strip_offset,
        offset.text = strip_text_offset,
        angle = find_firstlast("G", 3),
        hjust = find_firstlast("G", 4)
) +
#        geom_strip(
#        taxa1 = find_firstlast("U", 1),
#        taxa2 = find_firstlast("U", 2),
#        barsize = strip_barsize,
#        color = paletteer_d("ggsci::nrc_npg")[9],
#        label = "U",
#        fontsize = strip_fontsize,
#        offset = strip_offset,
#        offset.text = strip_text_offset,
#        angle = find_firstlast("U", 3),
#        hjust = find_firstlast("U", 4)
#) +
        theme(legend.position = "none")
```


```{r}
library(VennDiagram)

#cMake the plot
venn.diagram(
  x = list(
    genometa %>% filter(Collection =="BCO") %>% dplyr::select(ST) %>% unlist(), 
    genometa %>% filter(Collection =="AVC") %>% dplyr::select(ST) %>% unlist()
    ),
  category.names = c("BCO", "AVC"),
  filename = 'venn.png',
  output = TRUE ,
          imagetype="png" ,
          height = 480 , 
          width = 480 , 
          resolution = 300,
          compression = "lzw",
          lwd = 1,
          col=c("#440154ff", '#21908dff'),
          fill = c(alpha("#440154ff",0.3), alpha('#21908dff',0.3)),
          cex = 0.5,
          fontfamily = "sans",
          cat.cex = 0.3,
          cat.default.pos = "outer",
          cat.pos = c(-27, 27),
          cat.dist = c(0.055, 0.055),
          cat.fontfamily = "sans",
          cat.col = c("#440154ff", '#21908dff')
        )
```

```{r sankey_library, include=FALSE, echo=FALSE}
#Separate library calls to suppress the high charter message
library(highcharter)
library(htmlwidgets)
library(dplyr)
```

### Collection breakdown - HC5, Serovar, Source Niche and Source Details
```{r sankey_1, echo=FALSE}


sankey1 <- genometa %>% dplyr::select(Collection, ST_other, Clermont_type)

sankey1 <- apply(sankey1,
                 2,
                 FUN = function(x) paste0(x,
                                          " (",
                                          as.data.frame(table(x)[x])[,2],
                                          "/",
                                          as.character(nrow(metadata)),
                                          ")",
                                          " ",
                                          scales::percent(
                                                  as.data.frame(
                                                          table(x)[x])[,2]/nrow(metadata),
                                                  accuracy = 0.1
                                                  )
                                          )
                 )

sankey1 <- as.data.frame(sankey1)

sankey1_plot <- hchart(data_to_sankey(sankey1), "sankey", name = "Collection Breakdown")  %>%
        hc_plotOptions(series = list(dataLabels = list(style = list(fontSize = "15px", fontFamily = "Arial")))) %>% suppressWarnings()


```


```{r multi_site_stats, echo = TRUE}
library(ggpubr)

count_sites_by_ST <- genometa %>% filter(Collection == "BCO") %>% group_by(ST, Bird_Gross_Necropsy_ID) %>% summarise(counts = n()) %>% mutate(ST = paste0("ST", ST))

ST_by_clermontype <- genometa %>% group_by(ST, Clermont_type) %>% summarise(counts = n())

ST_by_clermontype <- ST_by_clermontype %>% 
        group_by(ST) %>%
        filter(counts == max(counts)) %>%
        arrange(ST,Clermont_type, counts) %>%
        dplyr::select(-counts) %>%
        mutate(ST = paste0("ST", ST))

count_sites_by_ST <- left_join(count_sites_by_ST, ST_by_clermontype) %>% arrange(match(count_sites_by_ST$ST, ST_by_clermontype$ST))

ST_order <- count_sites_by_ST %>% dplyr::select(ST, Clermont_type) %>% arrange(Clermont_type) %>% unique() %>% pull(ST)

count_sites_by_ST$ST <- factor(count_sites_by_ST$ST, levels = ST_order)

count_of_st <- genometa %>% filter(Collection == "BCO") %>% group_by(ST) %>% summarise(counts = n()) %>% mutate(ST = paste0("ST", ST)) %>% filter(counts > 3)

count_sites_by_ST <- count_sites_by_ST %>% filter(ST %in% count_of_st$ST)

count_sites_by_ST$Clermont_type <- factor(count_sites_by_ST$Clermont_type, levels = c("A","B1","B2","D","E", "G"))


p <- ggboxplot(count_sites_by_ST, x = "ST", y = "counts", color = "Clermont_type", palette = "jco",
          add = "jitter")


```


```{r process_vir}

vir_df <- BCO_APEC_simple_summary_N90L90 %>% dplyr::select(name, starts_with('vfdb'), starts_with('EC_'))

vir_df <- vir_df %>% column_to_rownames('name') %>% t() %>% as.data.frame()

vir_df <- vir_df %>% rownames_to_column('gene')

vir_df <- vir_df %>% mutate(gene = str_replace(gene, 'vfdb_|EC_custom_', ''))

vir_df <- vir_df %>% column_to_rownames('gene') %>% t()

vir_df[vir_df > 1] <- 1

vir_df <- as.data.frame(vir_df)

vir_df <- vir_df %>% dplyr::select(matches("astA"),
matches("cba"),
matches("celb"),
matches("cma"),
matches("cvaA"),
matches("cvaB"),
matches("cvaC"),
matches("cvi"),
matches("eilA"),
matches("eitA"),
matches("eitB"),
matches("eitC"),
matches("eitD"),
matches("etsA"),
matches("fimH"),
matches("fyuA"),
matches("gad"),
matches("gimB_marker"),
matches("hek"),
matches("hlyE"),
matches("hylF"),
matches("ibeA"),
matches("iha"),
matches("ipaH"),
matches("ireA"),
matches("iroN"),
matches("irp2"),
matches("iss"),
matches("iucD"),
matches("iutA"),
matches("kpsM"),
matches("kpsMT_II"),
matches("kpsMT_III"),
matches("lpfA"),
matches("malX"),
matches("mchB"),
matches("mchC"),
matches("mchF"),
matches("mcmA"),
matches("neuC"),
matches("ompT"),
matches("papA"),
matches("papB"),
matches("papC"),
matches("papD"),
matches("papE"),
matches("papF"),
matches("papGII"),
matches("papH"),
matches("papI"),
matches("papJ"),
matches("papK"),
matches("pic"),
matches("shiD"),
matches("sitA"),
matches("tia"),
matches("traT"),
matches("tsh"),
matches("usp"),
matches("vat"),
matches("yeeT")
) 

vir_df <- vir_df %>% 
        rowwise() %>% 
        mutate(fimH_new= sum(fimH, fimH_CP015076.1, na.rm = T)) %>% 
        dplyr::select(-fimH, -fimH_CP015076.1) %>%
        mutate(fyuA_new= sum(fyuA, fyuA_Z35106.1, na.rm = T)) %>% 
        dplyr::select(-fyuA, -fyuA_Z35106.1) %>%
        mutate(ibeA_new= sum(ibeA, ibeA_AY248744.1, na.rm = T)) %>% 
        dplyr::select(-ibeA, -ibeA_AY248744.1) %>%
        mutate(irp2_new = sum(irp2, `irp2_(gi:117624162)|VFG012550`, `irp2_(gi:218695586)|VFG034095`, `irp2_(gi:218705472)|VFG034101`, `irp2_(gi:222156713)|VFG034107`, `irp2_(gi:386629711)|VFG034099`, `irp2_(gi:387607663)|VFG034096`, `irp2_(gi:45441468)|VFG003119`, na.rm = T)) %>% 
        dplyr::select(-irp2, -`irp2_(gi:117624162)|VFG012550`, -`irp2_(gi:218695586)|VFG034095`, -`irp2_(gi:218705472)|VFG034101`, -`irp2_(gi:222156713)|VFG034107`, -`irp2_(gi:386629711)|VFG034099`, -`irp2_(gi:387607663)|VFG034096`, -`irp2_(gi:45441468)|VFG003119`) %>%
        mutate(iss_new= sum(`iss_type2_ETEC-B7a_NZ_AAJT00000000`, iss_type3_UTI89_CP000243, na.rm = T)) %>%            dplyr::select(-`iss_type2_ETEC-B7a_NZ_AAJT00000000`, -iss_type3_UTI89_CP000243) %>%
        mutate(iucD_new = sum(iucD, iucD_AY545598.5, na.rm = T)) %>%
        dplyr::select(-iucD, -iucD_AY545598.5) %>%
        mutate(iutA_new = sum(iutA, iutA_KU578032.1, na.rm = T)) %>%
        dplyr::select(-iutA, -iutA_KU578032.1) %>%
        mutate(`kpsMT(II)_new` = sum(`kpsMT(II)_K1:AE014075.1|CFT073|3527848-3529293`, `kpsMT(II)_K2:CP000468.1:|APEC-O1|3376659-3378102`, `kpsMT(II)_K5:CP002212.1|3385642-3387090`, na.rm = T)) %>%
        dplyr::select(-`kpsMT(II)_K1:AE014075.1|CFT073|3527848-3529293`, -`kpsMT(II)_K2:CP000468.1:|APEC-O1|3376659-3378102`, -`kpsMT(II)_K5:CP002212.1|3385642-3387090`) %>%
        mutate(neuC_new = sum(neuC, neuC_CP007275.1, na.rm = T)) %>%
        dplyr::select(-neuC, -neuC_CP007275.1) %>%
        mutate(ompT_new = sum(`ompT_chromosomal_O157-H7_NC_002695`, `ompT_episomal_HM210637.1`, na.rm = T)) %>%
        dplyr::select(-`ompT_chromosomal_O157-H7_NC_002695`, -`ompT_episomal_HM210637.1`) %>%
        dplyr::select(-papB, -papD, -papE, -papF, -papH, -papI, -papJ, -papK)

vir_df <- vir_df %>% mutate(sitA_new = sum(`sitA_(gi:218704670)|VFG034199`, `sitA_(gi:387606979)|VFG034194`, `sitA_(gi:387616520)|VFG034206`, `sitA_(gi:82776740)|VFG013039`,  `sitA_(gi:88770314)|VFG012575`,  `sitA_(gi:91210366)|VFG012573`, na.RM = T)) %>% dplyr::select(-`sitA_(gi:218704670)|VFG034199`, -`sitA_(gi:387606979)|VFG034194`, -`sitA_(gi:387616520)|VFG034206`, -`sitA_(gi:82776740)|VFG013039`,  -`sitA_(gi:88770314)|VFG012575`,  -`sitA_(gi:91210366)|VFG012573`)

vir_df <- vir_df %>% mutate(papC_new = sum(papC, papC_CU928161)) %>% dplyr::select(-papC, -papC_CU928161)

vir_df <- vir_df %>% mutate(tia_new = sum(tia_CP019944.1, tia_CU928161)) %>% dplyr::select(-tia_CP019944.1, -tia_CU928161)

vir_df <- vir_df %>% mutate(usp_new = sum(usp_AB056434, usp_AB056436, usp_AB056437, usp_AB056439)) %>% dplyr::select(-usp_AB056434, -usp_AB056436, -usp_AB056437, -usp_AB056439)

vir_df <- vir_df %>% rename_all(funs(stringr::str_replace_all(., '_.*', '')))
        
vir_df[vir_df > 1] <- 1

heatmap_vir <- pheatmap(vir_df, cluster_cols = TRUE)

vir_df <- vir_df %>% dplyr::select(all_of(heatmap_vir$tree_col$order))

binary_sum_perc_colnames <- function(df){
        df <- vir_df
        oldname <- colnames(df)
        n_w_gene <- colSums(df) %>% as.data.frame() %>% rownames_to_column('name') %>% rename('count' = '.')
        n_total <- nrow(df) %>% as.numeric()
        perc <- scales::percent(x = (n_w_gene$count/n_total), accuracy = 2)
        n_w_gene <- n_w_gene %>% mutate(count = as.character(count))
        n_total <- n_total %>% as.character(n_total)
        colnames(df) <- paste0(n_w_gene$name, " ", n_w_gene$count, "/", n_total, " (", perc, ")")
        return(df)
}

vir_df <- vir_df %>% binary_sum_perc_colnames()

vir_df <- vir_df %>% mutate(across(where(is.numeric), as.character))

vir_df <- vir_df %>% as.data.frame()

rownames(vir_df) <- BCO_APEC_simple_summary_N90L90$name



p2 <- gheatmap(
        p = p1_linear,
        data = vir_df,
        colnames_offset_y = -0.1,
        font.size = 3,
        hjust = 0,
        colnames = TRUE,
        colnames_position = "top",
        colnames_angle = 90,
        offset = 0.025,
        width = 15,
        color = rgb(220, 220, 220, max = 255, alpha = 50)
        ) +
theme(legend.position = "bottom",
        legend.box = "vertical",
        legend.key.size = unit(1, "mm"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8)
        ) +
scale_color_manual(
        values = c(mycolslist, c('0' = 'white', '1' = 'red')),
        na.value = 'grey',
        guide = "none"
        ) +
scale_fill_manual(
        values = c(mycolslist, c('0' = 'white', '1' = 'red')),
        na.value = 'grey',
        
        ) +
        ylim(NA, 350)




```

## IncF RST analysis
```{r}
pMLST <- read_delim("analysis/pMLST.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE,
    show_col_types = FALSE)

pMLST <- pMLST %>% pivot_wider(id_cols = 'name', names_from = 'pMLST_scheme', values_from = 'pMLST')

genometa <- left_join(genometa, pMLST)

long_inf_rst <- genometa %>% group_by(Collection, incf) %>% summarise(counts = n(), .groups = 'keep') %>% pivot_wider(id_cols = 'incf', names_from = 'Collection', values_from = 'counts') %>% mutate(across(everything(), ~replace_na(.,0)))

long_inf_rst <- long_inf_rst %>% mutate(total = AVC + BCO) %>% mutate(total = paste0(total,"/300 (", scales::percent(total/300, accuracy = 2),')'))


long_inf_rst <- long_inf_rst %>% mutate(AVC = paste0(AVC,"/95 (", scales::percent(AVC/95, accuracy = 2),')'))

long_inf_rst <- long_inf_rst %>% mutate(BCO = paste0(BCO,"/205 (", scales::percent(BCO/205, accuracy = 2),')'))

long_inf_rst<- long_inf_rst %>% mutate(incf = str_replace(incf, pattern = 'C4', replacement = 'F18'))


```



## ColV Stats

```{r}
contingency_table_ColV_Collection <- table(genometa$Collection, genometa$ColV)

small_subset_tree <- read.tree("analysis/trees/old/mashtree_QC_pass_AVC_BCO_bone_nondup.tree")

small_subset_tree$tip.label <- gsub("\\.R1$","", small_subset_tree$tip.label)

small_subset_genometa <- genometa %>% filter(name %in% small_subset_tree$tip.label)

table(small_subset_genometa$Collection, small_subset_genometa$ColV) %>% chisq.test()

chisq.test(contingency_table_ColV_Collection)
```

```{r write_metadata, echo=FALSE}



genometa <- genometa %>% mutate(Final_subset = if_else(name %in% small_subset_tree$tip.label, "Yes", "No"))

genometa <- genometa %>% mutate(Final_subset = if_else(grepl("AVC",name), "No", Final_subset))

#Write our metadata to file for our supplementary table
write_delim(genometa, "Figures/2023/Supplementary/Supplementary_Table_1.txt", delim = "\t")

```

## ST Table

```{r}

ST_table <- genometa %>% group_by(ST, Collection) %>% summarise(counts = n())

ST_table <- ST_table %>% pivot_wider(id_cols = 'ST', names_from = 'Collection', values_from = 'counts') %>% mutate(across(everything(), ~replace_na(.,0)))

ST_table <- ST_table %>% mutate(total = AVC + BCO) %>% arrange(desc(total)) %>% mutate(total = paste0(total,"/300 (", scales::percent(total/300, accuracy = 2),')'))


ST_table <- ST_table %>% mutate(AVC = paste0(AVC,"/95 (", scales::percent(AVC/95, accuracy = 2),')'))

ST_table <- ST_table %>% mutate(BCO = paste0(BCO,"/205 (", scales::percent(BCO/205, accuracy = 2),')'))

phylogroup_ST_intersection <- genometa %>% group_by(ST, Clermont_type) %>% summarise(clermontypes = n()) %>% mutate(type_and_count = paste0(Clermont_type, ' (',clermontypes,')')) %>% ungroup %>% group_by(ST) %>% mutate(Phylogroups = paste0(type_and_count, collapse = ",")) %>% dplyr::select(ST, Phylogroups) %>% unique()

ST_phylogroup_table <- inner_join(ST_table, phylogroup_ST_intersection, by = "ST") %>% dplyr::select(ST, Phylogroups, AVC, BCO, total)
```

## Virulence NMDS

```{r}
library(MASS)
library(vegan)
library(ellipse)
library(grid)
library(gridExtra)
library(dplyr)

#read data into variable data
mds.data <- genometa
rownames(mds.data) <- genometa %>% pull(name)
mds.metadata <- mds.data %>% dplyr::select(1:13, ST_other) %>% rename('Name' = 'name')

vir_df <- vir_df %>% mutate(across(where(is.character), as.numeric))

vir_df_MDS <- unique.data.frame(vir_df)

#VIRULENCE
#create distance object 
virdist <- vegdist(wisconsin(vir_df_MDS), method = "jaccard")
# Get the baseline solution: start with cmdscale
mds.null.vir <- isoMDS(virdist, tol=1e-7)
## See if you can get any better.
  repeat{
    mds.1.vir <- isoMDS(virdist, k=2, initMDS(virdist, k = 2), maxit=200, trace=FALSE, tol=1e-7)
    print(mds.1.vir$stress)
    print(mds.null.vir$stress)
    if(mds.1.vir$stress < mds.null.vir$stress) break
  }

# Scale solutions ("fix translation, rotation and scale")
mds.null.vir <- postMDS(mds.null.vir, virdist)
mds.1.vir <- postMDS(mds.1.vir, virdist)
# Compare solutions
plot(procrustes(mds.1.vir, mds.null.vir))
#convert to dataframe
mds.null.vir.df <- as.data.frame(mds.null.vir,
                             attr(x = mds.null.vir,
                                which = "dimnames"))
mds.1.vir.df <- as.data.frame(mds.1.vir, 
                          attr(x = mds.1.vir, 
                              which = "dimnames"))

mds.1.vir$points %>% rownames() -> mds.1.vir.df$Name
mds.1.vir.df <- left_join(mds.1.vir.df, mds.metadata)

mds.1.vir.df$ST <- as.factor(mds.1.vir.df$ST)

#Virulence by Sequence Type
b1 <- ggplot() +
  geom_point(data=mds.1.vir.df,
             aes(x=points.1,
                 y=points.2,
                 color = Collection), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=mds.1.vir.df$points.1,
                   y=mds.1.vir.df$points.2, 
                   color = mds.1.vir.df$Collection)) +
  scale_color_manual(breaks = collection_names, values = collection_colours)+
  ggtitle("Nonmetric MDS - Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal(xlim = c(-3, 4), ylim = c(-1.5, 2)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

b2 <- ggplot() +
  geom_point(data=mds.1.vir.df,
             aes(x=points.1,
                 y=points.2,
                 color = ST_other), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=mds.1.vir.df$points.1,
                   y=mds.1.vir.df$points.2, 
                   color = mds.1.vir.df$ST_other)) +
  scale_color_manual(breaks = ST_other_names, values = ST_other_cols)+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal(xlim = c(-3, 4), ylim = c(-1.5, 2)) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(b1, b2, nrow = 2)

```

## Resistance NMDS
```{r}

mds.data <- genometa
rownames(mds.data) <- genometa %>% pull(name)
mds.metadata <- mds.data %>% dplyr::select(1:13, ST_other, EC_custom_intI1_HQ730118.1) %>% rename('Name' = 'name', 'intI1 carriage' = EC_custom_intI1_HQ730118.1) %>% mutate(`intI1 carriage` = as.character(`intI1 carriage`))

#read data into variable data
card_df <- genometa %>% dplyr::select(name, starts_with("card_")) %>% column_to_rownames('name')

card_df_MDS <- unique.data.frame(card_df)

#VIRULENCE
#create distance object 
carddist <- vegdist(wisconsin(card_df_MDS), method = "jaccard")
# Get the baseline solution: start with cmdscale
mds.null.card <- isoMDS(carddist, tol=1e-7)
## See if you can get any better.
  repeat{
    mds.1.card <- isoMDS(carddist, k=2, initMDS(carddist, k = 2), maxit=200, trace=FALSE, tol=1e-7)
    print(mds.1.card$stress)
    print(mds.null.card$stress)
    if(mds.1.card$stress < mds.null.card$stress) break
  }

# Scale solutions ("fix translation, rotation and scale")
mds.null.card <- postMDS(mds.null.card, carddist)
mds.1.card <- postMDS(mds.1.card, carddist)
# Compare solutions
plot(procrustes(mds.1.card, mds.null.card))
#convert to dataframe
mds.null.card.df <- as.data.frame(mds.null.card,
                             attr(x = mds.null.card,
                                which = "dimnames"))
mds.1.card.df <- as.data.frame(mds.1.card, 
                          attr(x = mds.1.card, 
                              which = "dimnames"))

mds.1.card$points %>% rownames() -> mds.1.card.df$Name
mds.1.card.df <- left_join(mds.1.card.df, mds.metadata)

mds.1.card.df$ST <- as.factor(mds.1.card.df$ST)

#Virulence by Sequence Type
res_b1 <- ggplot() +
  geom_point(data=mds.1.card.df,
             aes(x=points.1,
                 y=points.2,
                 color = Collection), 
             size = 1) + # add the point markers
  stat_ellipse(aes(x=mds.1.card.df$points.1,
                   y=mds.1.card.df$points.2, 
                   color = mds.1.card.df$Collection)) +
  scale_color_manual(breaks = collection_names, values = collection_colours)+
  ggtitle("Nonmetric MDS - Resistance Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")

res_b2 <- ggplot() +
  geom_point(data=mds.1.card.df,
             aes(x=points.1,
                 y=points.2,
                 color = ST_other), 
             size = 1) + # add the point markers
  stat_ellipse(aes(x=mds.1.card.df$points.1,
                   y=mds.1.card.df$points.2, 
                   color = mds.1.card.df$ST_other)) +
  scale_color_manual(breaks = ST_other_names, values = ST_other_cols)+
  ggtitle("Nonmetric MDS - Resistance Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

res_b3 <- ggplot() +
  geom_point(data=mds.1.card.df,
             aes(x=points.1,
                 y=points.2,
                 color = `intI1 carriage`), 
             size = 1) + # add the point markers
  stat_ellipse(aes(x=mds.1.card.df$points.1,
                   y=mds.1.card.df$points.2, 
                   color = mds.1.card.df$`intI1 carriage`)) +
  scale_color_manual(breaks = c('0','1'), values = c('red', 'black'))+
  ggtitle("Nonmetric MDS - Resistance Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggpubr::ggarrange(res_b1, res_b2, res_b3, nrow = 1, legend = "bottom")

```

```{r}

fit.mds <- vegan::metaMDS(comm = dist(vir_df), distance = "jaccard")


data.scores <- as.data.frame(scores(fit.mds))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$name <- rownames(data.scores)  # create a column of site names, from the rownames of data.scores

data.scores <- genometa %>% dplyr::select(1:13, ST_other) %>% inner_join(data.scores)

head(data.scores) #look at the data

ggplot() +
  geom_point(data=data.scores,
             aes(x=NMDS1,
                 y=NMDS2,
                 color = ST_other), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=data.scores$NMDS1,
                   y=data.scores$NMDS2, 
                   color = data.scores$ST_other)) +
  scale_color_manual(breaks = ST_other_names, values = ST_other_cols)+
  ggtitle("Nonmetric MDS Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot() +
  geom_point(data=data.scores,
             aes(x=NMDS1,
                 y=NMDS2,
                 color = Collection), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=data.scores$NMDS1,
                   y=data.scores$NMDS2, 
                   color = data.scores$Collection)) +
  scale_color_manual(breaks = collection_names, values = collection_colours)+ 
  ggtitle("Nonmetric MDS Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5))
```




```{r}

fit.mds <- vegan::metaMDS(comm = dist(vir_df), distance = "jaccard")


data.scores <- as.data.frame(scores(fit.mds))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$name <- rownames(data.scores)  # create a column of site names, from the rownames of data.scores

data.scores <- genometa %>% dplyr::select(1:13, ST_other) %>% inner_join(data.scores)

head(data.scores) #look at the data

ggplot() +
  geom_point(data=data.scores,
             aes(x=NMDS1,
                 y=NMDS2,
                 color = ST_other), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=data.scores$NMDS1,
                   y=data.scores$NMDS2, 
                   color = data.scores$ST_other)) +
  scale_color_manual(breaks = ST_other_names, values = ST_other_cols)+
  ggtitle("Nonmetric MDS Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot() +
  geom_point(data=data.scores,
             aes(x=NMDS1,
                 y=NMDS2,
                 color = Collection), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=data.scores$NMDS1,
                   y=data.scores$NMDS2, 
                   color = data.scores$Collection)) +
  scale_color_manual(breaks = collection_names, values = collection_colours) +
  ggtitle("Nonmetric MDS Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}

#read data into variable data
mds.data <- genometa
rownames(mds.data) <- genometa %>% pull(name)
mds.metadata <- mds.data %>% dplyr::select(1:13, ST_other) %>% rename('Name' = 'name')

is_df <- genometa %>% dplyr::select(starts_with("ISfinder"))

rownames(is_df) <- genometa$name

is_df_MDS <- unique.data.frame(is_df)

#VIRULENCE
#create distance object 
ISdist <- vegdist(wisconsin(is_df_MDS), method = "jaccard")
# Get the baseline solution: start with cmdscale
mds.null.IS <- isoMDS(ISdist, tol=1e-7)
## See if you can get any better.
  repeat{
    mds.1.IS <- isoMDS(ISdist, k=2, initMDS(ISdist, k = 2), maxit=200, trace=FALSE, tol=1e-7)
    print(mds.1.IS$stress)
    print(mds.null.IS$stress)
    if(mds.1.IS$stress < mds.null.IS$stress) break
  }

# Scale solutions ("fix translation, rotation and scale")
mds.null.IS <- postMDS(mds.null.IS, ISdist)
mds.1.IS <- postMDS(mds.1.IS, ISdist)
# Compare solutions
plot(procrustes(mds.1.IS, mds.null.IS))
#convert to dataframe
mds.null.IS.df <- as.data.frame(mds.null.IS,
                             attr(x = mds.null.IS,
                                which = "dimnames"))
mds.1.IS.df <- as.data.frame(mds.1.IS, 
                          attr(x = mds.1.IS, 
                              which = "dimnames"))

mds.1.IS$points %>% rownames() -> mds.1.IS.df$Name
mds.1.IS.df <- left_join(mds.1.IS.df, mds.metadata)

mds.1.IS.df$ST <- as.factor(mds.1.IS.df$ST)

#Virulence by Sequence Type
b1 <- ggplot() +
  geom_point(data=mds.1.IS.df,
             aes(x=points.1,
                 y=points.2,
                 color = Collection), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=mds.1.IS.df$points.1,
                   y=mds.1.IS.df$points.2, 
                   color = mds.1.IS.df$Collection)) +
  scale_color_manual(breaks = collection_names, values = collection_colours)+
  ggtitle("Nonmetric MDS Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5))

b2 <- ggplot() +
  geom_point(data=mds.1.IS.df,
             aes(x=points.1,
                 y=points.2,
                 color = ST_other), 
             size = 1.5) + # add the point markers
  stat_ellipse(aes(x=mds.1.IS.df$points.1,
                   y=mds.1.IS.df$points.2, 
                   color = mds.1.IS.df$ST_other)) +
  scale_color_manual(breaks = ST_other_names, values = ST_other_cols)+
  ggtitle("Nonmetric MDS Virulence Profile")+
  xlab("Dimension 1")+
  ylab("Dimension 2")+
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5))
```


